---
title: "Midterm - Data Wrangling - Osman Khan"
output:
  pdf_document: default
  html_notebook: default
---

###Question 1
```{r include = FALSE}

## import data, libraries
library("RODBC")
library(dplyr)
library(tidyverse)
library(lubridate)
library(stringr)
library(Hmisc)
library(SASxport)
library(sqldf)
initdat<-read.xport("\\\\Mac/Home/Downloads/DIQ_I.XPT")

```

On a high level, I examined whether the data was intact, missingness and counts.  

Most obviously, we expect the count to be 9575. A quick check confirms that it is `r count(initdat)`.

As SQL commands are more intuitive to me, I conducted my high-level analyses in a mixture of SQL and R.

Using SQL, we can also quickly examine the missingess, and find it to be 
`r percent_missing = sum(is.na.data.frame(initdat)) * 100 / (dim(initdat)[1] * dim(initdat)[2])`

Next, I wanted to check on a column by column basis if I had the same counts as three random factor columns.  
As seen below, it turned out it did.
```{r}
count_check <- function(df_col, counts) {
  discrete_sums <- c(sum(df_col ==  1, na.rm = TRUE),
  sum(df_col ==  2, na.rm = TRUE),
  sum(df_col ==  3, na.rm = TRUE),
  sum(df_col ==  7, na.rm = TRUE),
  sum(df_col ==  9, na.rm = TRUE),
  sum(is.na(df_col)))
  print(discrete_sums)
  return (discrete_sums == counts)
}

# counts taken from online data description at 
# https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DIQ_I.htm#Data_Processing_and_Editing

count_check(initdat$DIQ010, c(856, 8568, 147, 0, 4, 0))

count_check(initdat$DIQ160, c(513, 5521, 0, 0, 11, 3530))

count_check(initdat$DIQ080, c(194, 644, 0, 0, 8, 8729))


```
I made a brief data visualization of the column by column missingness that can be seen below:
```{r include = FALSE}


```

> Issues with Dataset

* Majority of columns should be factors {1, 2, 7, and 9 (yes, no, don't know and refused) do not have a meaningful mathematical relationship, but attempting to linearly model this dataset would result in R mistaking them for having a mathematical hierarchy.}
  + These columns need to be factorized.

* Many numeric columns have missing-value stand-ins like 7777 or 9999 that would horribly distort any findings.
  + Example: how many months has it been since insulin was taken might have a ridiculously high mean because of 999 and 7777 skewing the results.
  + These have to be removed.

* In three instances, a column of numerical time values is followed by a second column of corresponding units of time.
  + Example: 1 | 2 refers to 1 month, but 3 | 1 might refer to 3 years This is unwieldy as analyses of the first column will fail to take the unit of time into account.
  + I removed the second columns and instead established a single column of rates.

```{r}
plot(cars)
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
